\section{The Library Libhomology}
\label{program:libhomology}

With {\bf libhomology} we provide an expandable framework for all kinds of homology computations.
In our context, it is used as the foundation of the program kappa (see Section \ref{chapter_program:kappa}).
In the following we explain use and essential details of its classes.

The template class \progclass{ChainComplex} is the core of {\bf libhomology}.
Here we think of a chain complex as a finite sequence of compatible matrices $D_n$ satisfying $D_{n-1} D_n = 0$ which we call differential.
We do not mention bases.
Given a \progclass{ChainComplex}, our goal is to determine its homology.
Therefore one needs an implementation of the coefficient ring \progclass{CoefficientT} (see Subsection \ref{program:libhomology:CoefficientT}) and the matrix type \progclass{MatrixT} (see Subsection \ref{program:libhomology:MatrixT}) of the differentials.
The class \progclass{HomologyT} (see Subsection \ref{program:libhomology:HomologyT}) specifies the scope of homological information one wants to extract.
These should be derived using the class \progclass{DiagonalizerT} (see Subsection \ref{program:libhomology:DiagonalizerT}) which diagonalizes matrices by applying row or column operations.

The implementations of the classes \progclass{ChainComplex}, \progclass{MatrixT}, \progclass{DiagonalizerT} and \progclass{HomologyT} are interdependent,
so we recommend to skim over the details on the first reading.

\input{chapter_program/subsection_chaincomplex_type}
\input{chapter_program/subsection_coefficient_type}
\input{chapter_program/subsection_matrix_type}
\input{chapter_program/subsection_diagonalizer_type}
\input{chapter_program/subsection_homology_type}